---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## load the FAS litters data
```{r}
litters_df=read_csv("./data/FAS_litters.csv")
litters_df=janitor::clean_names(litters_df)
```

##`select` what we want to keepm select function operates on columns 

choose some columns and not others

```{r}
select(litters_df,group,litter_number)
```

```{r}
select(litters_df,group,gd0_weight)
```

## we will choose columns from gd0_weight to gd_of_birth(all between)
```{r}
select(litters_df,group,gd0_weight:gd_of_birth)
```


## select what columns we want to lose
```{r}
select(litters_df,-litter_number)
```

##renaming columns, using select function, we get to keep what columns we want and rename them
```{r}
select(litters_df,GROUP=group,LITTer_NUmber=litter_number)
```

##in dplyer there is a delicated function used for renaming, rename function, this way we do not delete any columns, just rename
```{r}
rename(litters_df,GROUP=group,LITTer_NUmber=litter_number)
```


?select_helpers we are interested in funtions that let us select colums that i want to keep based on some criteria

```{r}
select(litters_df, starts_with("gd"))
```


##we move litter_number to the first column and keep everything else
```{r}
select(litters_df, litter_number,everything())
```

##use relacate function to relocate litter_number to first column
```{r}
relocate(litters_df,litter_number)
```


##filter function operates on rows

```{r}
filter(litters_df,gd0_weight < 22)
```

```{r}
filter(litters_df,gd0_weight >= 22)
```


##logical operator we want to ask if gd_of_birth  equals to 20, == tests for equality, we are not giving gd_of_birth a value of 20, just test if gd_of_birth  equals to 22 is true or not
```{r}
filter(litters_df,gd_of_birth == 20)
```


##we want to see all the rows that gd_of_birth  equals to 22 is not true
```{r}
filter(litters_df,!(gd_of_birth == 20))
```
the code chunk below will give me everything that gd_of_birth not equals 20
another way to do 

```{r}
filter(litters_df,gd_of_birth != 20)
```

## i want both to be true
```{r}
filter(litters_df,gd0_weight>=22,gd_of_birth == 20)
```

```{r}
filter(litters_df,group=="Con7")
```

```{r}
filter(litters_df,group=="Mod8")
```

## want to see if group equals to either Con7 or Mod8

```{r}
filter(litters_df,group %in% c("Con7","Mod8"))
```



















