---
title: "Data import"
output: github_document
---

```{r setup}
library(tidyverse)
```

the shortcut for r code chunk is option+command+i

##Read in some data

Read in the litters dataset

printing in the console
```{r}
litters_df=read_csv("./data/FAS_litters.csv")
litters_df=janitor::clean_names(litters_df)
```
after the clean names everycolumn name will be displayed as snake case

##take a look at the data 
```{r}
litters_df
```

Only gives the first 6 rows,not used often
```{r}
head(litters_df)
```


check the last few rows in our dataset
```{r}
tail(litters_df)
```

the signal :: means I will use skim function in the skimr package
```{r}
skimr::skim(litters_df)
```

##options to read csv

```{r}
litters_df=read_csv("./data/FAS_litters.csv",skip=10, col_names=FALSE)
```

check out`read_csv(` for more information


##other file formats,excel and sas

read in an excel file
```{r}
library(readxl)
mlb_df=read_excel("./data/mlb11.xlsx")
mlb_df=read_excel("./data/mlb11.xlsx", range= "A1:F7")
mlb_df
```


READ in a SAS file

```{r}
library(haven)
pulse_df=read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```










